Calculating Entropies With Nested Sampling

### 2017-03-05 paper/ms.tex
More results

A purely sinusoidal signal has the form
\begin{align}
y(t) &= A \sin \left(\frac{2\pi t}{T} + \phi\right)
\end{align}
where $A$ is the amplitude, $T$ is the period, and $\phi$ is the phase.
Throughout this section, I parameterise the period by its logarithm,
$\tau = \log_{10} T$. I assumed the following priors:
\begin{align}
\ln A   &\sim \textnormal{Normal}\left(0, 0.1^2\right)  \\
\tau    &\sim \textnormal{Uniform}(-1, 0)  \\
\phi    &\sim \textnormal{Uniform}(0, 2\pi)
\end{align}
and the following conditional prior for the data:
\begin{align}
d_i | A, T, \phi &\sim \textnormal{Normal}\left(y(t_i), 0.1^2\right).
\end{align}
That is, the data is just the signal $y(t)$ observed at particular times
$\{t_1, ..., t_n\}$, with gaussian noise of standard deviation 0.1.
The amplitude of the sinusoid is very likely to be around 10 times the
noise level, and the period is between 0.1 and 1 times the duration of
the data.

Letting $\tau = \log_{10} T$, and treating
$A$ and $\phi$ are nuisance parameters, I calculated the entropies
of $\tau$, $\d$, and $(\tau, \d)$.
These results were computed using a tolerance of 0.001 for each of the three
runs. The results are:
\begin{align}
H_{\tau}    &= 7.307 \pm 0.027 \\
H_{\d}      &= 900.6 \pm 1.1   \\
H_{\tau,\d} &= 882.2 \pm 1.1
\end{align}

@@

\subsection{Uneven observations}

The results for the uneven observing schedule are as follows:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 106.8 \pm 1.1   \\
H'_{\tau, \d} &= 97.0 \pm 1.0    \\
I_{\tau, \d} &= 10.9 \pm 0.4
\end{align}
which appears to be higher than under the even observing schedule,
though the uncertainties
are not small enough to be certain of this.

An example signal observed with the even and uneven observing schedules
is shown in Figure~\ref{fig:sinewave}.

>>>>

A purely sinusoidal signal has the form
\begin{align}
y(t) &= A \sin \left(\frac{2\pi t}{T} + \phi\right)
\end{align}
where $A$ is the amplitude, $T$ is the period, and $\phi$ is the phase.
Throughout this section, I parameterise the period by its logarithm,
$\tau = \log_{10} T$. I assumed the following priors:
\begin{align}
\ln A   &\sim \textnormal{Normal}\left(0, 0.1^2\right)  \\
\tau    &\sim \textnormal{Uniform}(-1, 0)  \\
\phi    &\sim \textnormal{Uniform}(0, 2\pi)
\end{align}
and the following conditional prior for the data:
\begin{align}
d_i | A, T, \phi &\sim \textnormal{Normal}\left(y(t_i), 0.1^2\right).
\end{align}
That is, the data is just the signal $y(t)$ observed at particular times
$\{t_1, ..., t_n\}$, with gaussian noise of standard deviation 0.1.
The amplitude of the sinusoid is very likely to be around 10 times the
noise level, and the period is between 0.1 and 1 times the duration of
the data. An example signal observed with the even and uneven observing schedules
is shown in Figure~\ref{fig:sinewave}.

\subsection{Results}
Letting $\tau = \log_{10} T$, and treating
$A$ and $\phi$ are nuisance parameters, I calculated the entropies
of $\tau$, $\d$, and $(\tau, \d)$.
These results were computed using a tolerance of 0.001 for each of the three
runs. The results are:
\begin{align}
H_{\tau}    &= 7.307 \pm 0.027 \\
H_{\d}      &= 900.6 \pm 1.1   \\
H_{\tau,\d} &= 882.2 \pm 1.1
\end{align}

@@

The results for the uneven observing schedule were:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 895.5 \pm 1.5   \\
H'_{\tau, \d} &= 874.7 \pm 1.5    \\
I_{\tau, \d} &= 30.13 \pm 0.70
\end{align}

The difference in mutual informations between the two observation
schedules is $2.3 \pm 1.3$ nats, suggesting a slight advantage to the
uneven observing schedule, in terms of the data's relevance to inference
of the log-period of the signal.
In a serious experimental design situation, the relevance is not the only
consideration (or the answer to every experimental design problem would
be to obtain more data without limit), but it is a very important one.
